### 2025-04-07 02_models_1.Rmd
added creating folders, runs models with all species and only with temp

---
title: "02_models_1"
author: "Sophie Moreau"
date: "2024-09-27"
output: html_document
editor_options: 
  chunk_output_type: console
---
#General Additive Models
Running the predictions for all GAMs with the smaller data frames to reduce running 
time.

```{r}
# packages relevant for data analysis
library(tidyverse)
library(here)
library(readr)
library(readxl)
library(gratia)
library(mgcv)
library(forcats)
library(ggpubr)
library(RColorBrewer)
library(broom)
library(grid)
library(DHARMa)
library(mgcViz)
library(gridExtra)
library(grid)
library(viridis)

# function for the temperature only GAMs
source(here("function_GAM.R"))

# function for temp and depth models 
source(here("function_temp_depth.R"))
```